---
title: "R Basics Day 1"
author: "Pietro Franceschi"
date: "9/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Outline

## Rmd Files

In Rmd the chunks of R are represented by shaded areas

```{r}
# create a variable called a which cointains a sequence of 10 numbers
a <- 1:10
```

## Packages

Packages are sets of functions (data) which are used to extend the potential of R. 
Packages are community maintained.

Reliable sources of packages are

* CRAN
* bioconductor
* github

## Data Types

```{r}
## Basic data types

num <- 1.4     ## numbers
pippo <- "1.4" ## string
bol <- TRUE  ## boolean

```


## Data Frames and tibbles
DThe data frame is the analogous of an excel table. It can contain columns of differnt type, but with the same length

For us `tibbles` are a new version of `data.frames`

 
##  Extracting rows and columns

Battleship approach. Use square brackets [row,col]


```{r}
## Load the iris dataset
data(iris)


## Extract one column (the first)
iris[,1]

## Extract the first row 
iris[1,]

## Extract the first and the third element of the second column
iris[c(1,3),2]

## Extract elements by name. First element of the "Species" column

iris[1,"Species"]

## Fast way to get a column if you are dealing with a df
iris$Species ## get out only the "Species" column



```


## Loading data!

The easiest way is to use the button "Import Dataset" in the "environment"
tab

```{r}
## load the data by using the utilities provided in readr
library(readr)
skulls <- read_csv("skulls.csv")

```


## Adding Columns

```{r}
## add to the iris df one column containing the ratio between sepal and petal

iris$myratio <- iris$Sepal.Length/iris$Sepal.Width

```





* DIY - Practical
    * Subsetting
    * Creating new columns
    
    
    
    
* R Functions
* DIY - Practical
* ggplotting, first steps
* DIY - practical


